TEMPERATURES AND METALLICITIES OF M DWARFS IN THE APOGEE SURVEY

### 2019-08-21 paper/draft.tex
report consistent scatters

The data-driven approach of \thecannon\ is ideal in certain cases:
if stellar labels are known for a small number of stars
but there are spectra taken for many more;
if it is computationally expensive to obtain labels for a star, and
there are many stars that need labels;
or if there are spectral models or techniques that work in one wavelength range or resolution but not in another.
Existing methods to model M dwarf spectra in the NIR at high resolution are computationally expensive, calibrated over a narrow range of $\teff$ and metallicity, and/or focused on the $J$-band \citep{Lindgren:2016, Lindgren:2017}. \thecannon\ thus fills this niche: it doesn't require using specific lines or opacity information that may be missing from the models; instead it allows us to determine labels from a lot of low level metallicity information present in thousands of lines, and as we demonstrate, does so with very good precision.

@@

While a detailed analysis of atmospheric model limitations is beyond the scope of this paper, 
our results provide an avenue to compare the FGK star metallicity scale 
to the less well-understood M dwarf metallicity scale.
These results find that atmospheric metallicities are systematically metal-poor biased
compared to \cannon-based metallicities trained on sources with metallicities calibrated
to those of FGK companions. 
At the high temperature end ($\teff > 3600$\,K), the ASPCAP metallicity bias is $\sim-0.15\,$dex with a
scatter of $\sim0.06\,$dex relative to \cannon\ metallicities, and increases to a bias of $\>-0.26\,$dex 
and scatter of $\>0.1\,$dex at the lower temperature end ($\teff < 3300$\,K) (Figure \ref{fig:met_comp}).
We suspect that this metal-poor bias, while not explored to a great extent in this work, is due 
to the line lists of the models: 
an effect in which the optimizer of the pipeline may be lowering the metallicity of the fit 
to compensate for the missing lines or opacities.
Further analysis would need to be done to determine if the metallicity bias
is found in other model grids (besides the ATLAS/MARCS models used by the ASPCAP pipeline), 
and if the effect is present at other wavelengths.

>>>>

The data-driven approach of \thecannon\ is ideal in certain cases:
if stellar labels are known for a small number of stars
but there are spectra taken for many more;
if it is computationally expensive to obtain labels for a star, and
there are many stars that need labels;
or if there are spectral models or techniques that work in one wavelength range or resolution but not in another.
Existing methods to model M dwarf spectra in the NIR at high resolution are computationally expensive, often calibrated over a narrow range of Teff and/or metallicity. \thecannon\ thus fills this niche: it doesn't require using specific lines or opacity information that may be missing from the models; instead it allows us to determine labels from a lot of low level metallicity information present in thousands of lines, and as we demonstrate, does so with very good precision.

@@

While a detailed analysis of atmospheric model limitations is beyond the scope of this paper, 
our results provide an avenue to compare the FGK star metallicity scale 
to the less well-understood M dwarf metallicity scale.
These results find that atmospheric metallicities are systematically metal-poor biased
compared to \cannon-based metallicities trained on sources with metallicities calibrated
to those of FGK companions. 
At the high temperature end ($\teff > 3600$\,K), the ASPCAP metallicity bias is $-0.12-0.15\,$dex with a scatter of $0.05-0.06\,$dex relative to \cannon\ metallicities, and increases to a bias of $-0.3\,$dex and scatter of $0.13\,$dex at the lower temperature end ($\teff < 3200$\,K) (Figure \ref{fig:met_comp}).
We suspect that this metal-poor bias, while not explored to a great extent in this work, is due 
to the line lists of the models: 
an effect in which the optimizer of the pipeline may be lowering the continuum level and metallicity of the fit to compensate for the missing lines or opacities.
Further analysis would need to be done to determine if the metallicity bias
is found in other model grids (besides the ATLAS/MARCS models used by the ASPCAP pipeline), 
and if the effect is present at other wavelengths.
